---
title: "Albert Yen Analysis"
output: html_notebook
---


Loading Libraries and Data
```{r}
library(sf)
library(ggplot2)
library(tmap)
library(tmaptools)
library(leaflet)
library(dplyr)
options(scipen = 999)
library(readxl)
library(gridExtra)
input = as.data.frame(read_excel("C:/Users/alber/Downloads/Stat382/Project4/Cleaned_Data.xlsx"))
input[,c(24:27)] = lapply(input[,c(24:27)],as.factor)
```

#Exploratory Analysis

###Vote Share by urbanicity and Atlanta
```{r}
Total = tapply(input$Total_votes, input$urbanicity,sum)
B_vote = tapply(input$Total_votes*input$BidenVoteShare/100, input$urbanicity,sum)
#urbanicity = c("Rural","Suburban", "Urban")
voters = tapply(input$`TOTAL VOTERS`, input$urbanicity,sum)
par(mfrow = c(2,2))

barplot(B_vote/Total, col="light blue", main = "Biden's Vote Share",ylim=c(0,0.8))
barplot(voters/sum(voters), col="light blue", main = "Share of Registered Voters", ylim=c(0,0.5))

B_atlanta_votes = tapply(input$Total_votes*input$BidenVoteShare/100, input$in_atlanta_metro,sum)
atlanta_voters = tapply(input$`TOTAL VOTERS`, input$in_atlanta_metro,sum)
atlanta_voted = tapply(input$Total_votes, input$in_atlanta_metro,sum)

barplot(B_atlanta_votes/atlanta_voted, col="light blue", main = "Biden's Vote Share", names.arg =c( "Outside of Atlanta", "Atlanta"), ylim=c(0,0.7))
barplot(atlanta_voted/sum(atlanta_voted), col="light blue", names.arg =c("Outside of Atlanta", "Atlanta"), main = "Share of Registered Voters", ylim=c(0,0.6))
```
###Margin Increases by Urbanicity and Atlanta Metro
```{r}
votes2016 = tapply(input$Total_votes*(100-input$voteChange)/100, input$urbanicity,sum)
new_margin = tapply(input$new_margin*input$Total_votes/100, input$urbanicity, sum)/Total
old_margin = tapply(input$old_margin*input$Total_votes*(100-input$voteChange)/100/100, input$urbanicity, sum)/votes2016

par(mfrow = c(1,2))
barplot(new_margin-old_margin,ylim=c(-0.05,0.1), main="Change in Margin of Victory", col="light blue")

atlanta_votes2016 = tapply(input$Total_votes*(100-input$voteChange)/100, input$in_atlanta_metro,sum)
atlanta_new_margin = tapply(input$new_margin*input$Total_votes/100, input$in_atlanta_metro, sum)/atlanta_voted
atlanta_old_margin = tapply(input$old_margin*input$Total_votes*(100-input$voteChange)/100/100, input$in_atlanta_metro, sum)/atlanta_votes2016
barplot(atlanta_new_margin-atlanta_old_margin,ylim=c(0,0.1), main="Change in Margin of Victory", col="light blue", names.arg =c("Outside of Atlanta", "Atlanta"))
```
###Democrat's margin increase by urbanicity, Atlanta, and 2016 results
```{r}
winner2016_2016voted = tapply(input$Total_votes*(100-input$voteChange)/100, input$past_result,sum)
winner2016_voted = tapply(input$Total_votes, input$past_result,sum)
winner2016_voter = tapply(input$`TOTAL VOTERS`, input$past_result,sum)
winner2016_new_margin = tapply(input$new_margin*input$Total_votes/100, input$past_result, sum)/winner2016_voted
winner2016_old_margin = tapply(input$old_margin*input$Total_votes*(100-input$voteChange)/100/100, input$past_result, sum)/winner2016_2016voted

par(mfrow = c(2,2))
barplot(new_margin-old_margin,ylim=c(-0.05,0.1), main="Change in Margin of Victory", col="light blue")
barplot(atlanta_new_margin-atlanta_old_margin,ylim=c(0,0.1), main="Change in Margin of Victory", col="light blue", names.arg =c("Outside of Atlanta", "Atlanta"))
barplot(winner2016_new_margin-winner2016_old_margin,ylim=c(0,0.1), main="Change in Margin of Victory", col="light blue", names.arg = c("Clinton Counties", "Trump Counties"))
```


###Turnout by by urbanicity, Atlanta, and 2016 results
```{r}
par(mfrow = c(2,2))
turnout = sum(Total)/sum(voters)
barplot(Total/voters, col="light blue", main = "Turnout",
        width = c(0.1,0.1,0.1))
barplot(atlanta_voted/atlanta_voters,main = "Turnout", col="light blue", ylim=c(0,0.7), names.arg =c("Outside of Atlanta", "Atlanta"))
barplot(winner2016_voted/winner2016_voter,main = "Turnout", col="light blue", ylim=c(0,0.7), names.arg =c("Clinton Counties", "Trump Counties"))
```
###Voter turnout Increases by urbanicity, Atlanta, and 2016 results
```{r}
winner2016_2016total_voters = tapply(input$Total_votes*(100-input$voteChange)/100/(input$Turnout-input$TurnoutChange)*100, input$past_result,sum)

atlanta_2016total_voters = tapply(input$Total_votes*(100-input$voteChange)/100/(input$Turnout-input$TurnoutChange)*100, input$in_atlanta_metro,sum)

total_voters2016 = tapply(input$Total_votes*(100-input$voteChange)/100/(input$Turnout-input$TurnoutChange)*100, input$urbanicity,sum)

winner2016_2016turnout = winner2016_2016voted/winner2016_2016total_voters 
winner2016_turnout = winner2016_voted/winner2016_voter

turnout2016 = votes2016/total_voters2016
increase = Total/voters-turnout2016

atlanta_turnout2016 = atlanta_votes2016/atlanta_2016total_voters
atlanta_turnout_increase = atlanta_voted/atlanta_voters - atlanta_turnout2016

par(mfrow = c(2,2))
barplot(increase,main = "Turnout Increase", col="light blue", ylim=c(0,0.08))

barplot(winner2016_turnout-winner2016_2016turnout,main = "Turnout Increase", col="light blue", ylim=c(0,0.08), names.arg =c("Clinton Counties", "Trump Counties"))

barplot(atlanta_turnout_increase,main = "Turnout Increase", col="light blue", ylim=c(0,0.08), names.arg =c("Outside of Atlanta", "Atlanta"))
```
###Change in total number of votes cast by urbanicity, Atlanta, and 2016 results
```{r}
input$voted2016 = round(input$Total_votes*(100-input$voteChange)/100)
input$raw_vote_increase = input$Total_votes-input$voted2016

par(mfrow = c(2,2))

barplot(tapply(input$raw_vote_increase, input$urbanicity, sum)/as.vector(tapply(input$voted2016,input$urbanicity,sum)), col="light blue", main="Votes Cast Increase", ylim=c(0,.35))

barplot(tapply(input$raw_vote_increase, input$in_atlanta_metro, sum)/as.vector(tapply(input$voted2016,input$in_atlanta_metro,sum)), col="light blue", main="Votes Cast Increase", ylim=c(0,.35),names.arg =c("Outside of Atlanta", "Atlanta") )

barplot(tapply(input$raw_vote_increase, input$past_result, sum)/as.vector(tapply(input$voted2016,input$past_result,sum)), col="light blue", main="Votes Cast Increase", ylim=c(0,.35), names.arg =c("Clinton Counties", "Trump Counties") )


```
#Generating Map Graphics

###read in map data
```{r}
mymap <- st_read("C:/Users/alber/Downloads/Stat382/Project4/georgia/Georgia_Counties.shp", stringsAsFactors=F)
names(mymap)[names(mymap)=="Name"] ="County"
map_and_data = inner_join(mymap,input)
```
###By Margin
```{r}
map = ggplot(map_and_data)
map+geom_sf(aes(fill=new_margin))+scale_fill_gradient2(midpoint = 0, low = "#FF0000", mid = "white", high = "#0000ff", space = "Lab", name="New Margin" )+
   labs(title = "Democrat's Margin by County, 2020") +theme(title =element_text(size=18, face='bold'))


```
###Margin Change
```{r}
map+geom_sf(aes(fill=margin_change))+scale_fill_gradient2(midpoint = 0, low = "#FF0000", mid = "white", high = "#0000ff", space = "Lab", name="New Margin" )+
   labs(title = "Democrat's Margin Change, 2016-2020") +theme(title =element_text(size=18, face='bold'))
```
###Share of Senior Voters
```{r}
o=map+geom_sf(aes(fill=senior))+scale_fill_gradient(low= 'white',high='#FF0000', space = "Lab", name="Percentage")+
   labs(title = "Share of Electorate over 65") +theme(title =element_text(size=18, face='bold'))
y=map+geom_sf(aes(fill=youth))+scale_fill_gradient(low= 'white',high='#0000ff', space = "Lab", name="Percentage")+
   labs(title = "Share of Electorate under 30") +theme(title =element_text(size=18, face='bold'))
grid.arrange(y,o, ncol=2)
```

###Demographic Overview
```{r}
w = map+geom_sf(aes(fill=WH))+scale_fill_gradient(low= 'white',high='#FF0000', space = "Lab", name="Percentage")+
   labs(title = "Share of White Voters") +theme(title =element_text(size=18, face='bold'))
b = map+geom_sf(aes(fill=BH))+scale_fill_gradient(low= 'white',high='#0000ff', space = "Lab", name="Percentage")+
   labs(title = "Share of Black Voters") +theme(title =element_text(size=18, face='bold'))
h = map+geom_sf(aes(fill=HP))+scale_fill_gradient(low= 'white',high='green', space = "Lab", name="Percentage")+
   labs(title = "Share of Hispanic Voters") +theme(title =element_text(size=18, face='bold'))
a = map+geom_sf(aes(fill=AP))+scale_fill_gradient(low= 'white',high='orange', space = "Lab", name="Percentage")+
   labs(title = "Share of Asian Voters") +theme(title =element_text(size=18, face='bold'))
grid.arrange(w, b, a, h, ncol=2, nrow=2)

```

