---
title: "Stat 382 Final Project"
output: html_notebook
---
#Data Cleaning

####Variable Meaning
AI = Native American, percentage of registered voters
AP = Asian and Pacific Islander, percentage of registered voters
BH = Black, percentage of registered voters
HP = Hisanpic, percentage of registered voters
WH = White, percentage of registered voters
OT = Other, percentage of registered voters
UK = race unknown, percentage of registered voters
Total_votes = Total votes casted
TOTAL VOTERS = Total number of registered voters
voteChange = Change in total number of votes casted, compared with 2016
Education = Percentage of people aged 25 and older who has a bachelor's degree
pop_density = people per square mile
new_margin = Democrat's margin of victory in 2020 in percent
old_margin = Democrat's margin of victory in 2016 in percent
margin_change = margin change between 2016 and 2020
current_result = Which party won in 2020
past_result = which party won in 2016
senior = percentage of registered voter who are over 65 years old
youth = percentage of registered voter who are under 30 years old
in_atlanta_metro = Is the county part of the Atlanta metropolitan area or not

###Loading Election and Electorate Data
```{r}
library(readxl)
GAelection = as.data.frame(read_excel("C:/Users/alber/Downloads/Stat382/Project4/Georgia.xlsx"))[-160,]
GAvoter =  as.data.frame(read_excel("C:/Users/alber/Downloads/Stat382/Project4/Active_Voters_by_Race_Gender_as_of_November_1_2020.xlsx"))[-160,c(-1:-23)]
GAvoter[,-1] = sapply(GAvoter[,-1], function(x) return(round(x*100/GAvoter$`TOTAL VOTERS`,2)))
ga2016 = as.data.frame(read_excel("C:/Users/alber/Downloads/Stat382/Project4/GA2016.xlsx"))
GA = cbind(GAelection, GAvoter)
BidenVoteShare = round(100*GA$`B_Total Votes`/GA$Total, 2)
TrumpVoteShare = round(100*GA$`T_Total Votes`/GA$Total, 2)
GA$BidenVoteShare = BidenVoteShare
GA$TrumpVoteShare = TrumpVoteShare

data = GA[,c(1,17:27)]
data$Turnout = round(100*data$Total/data$`TOTAL VOTERS`,2)
data$voteChange = round((100*data$Total/ga2016$Total)-100,2)
turnout2016 = as.data.frame(read_excel("C:/Users/alber/Downloads/Stat382/Project4/gaTurnout2016.xlsx"))
data$TurnoutChange = data$Turnout-turnout2016$Turnout
names(data)[names(data) == "Total"] = "Total_votes"
```
###Add in population density and income
```{r}
pop = read.csv("C:/Users/alber/Downloads/Stat382/Project4/LandArea.csv")[,c(2:4)]
pop$Population = as.numeric(pop$Population)
pop$pop_density=pop$Population/pop$Land.Area
data=merge(data,pop)
data=subset(merge(pop,data), select=-c(Population,Land.Area))
income = as.data.frame(read_excel("C:/Users/alber/Downloads/Stat382/Project4/incomeED.xlsx"))[,c(-2,-4,-5,-6)]
data=merge(data,income)
```
###Add in percentage of seniors and youth
```{r}
age = as.data.frame(read_excel("C:/Users/alber/Downloads/Stat382/Project4/Active_Voters_by_Race_Gender_and_Age_as_of_November_1_2020.xlsx"))[c(seq(1,1753,by=11)),c(-1:-25)][-160,]
data = cbind(data,age[,c(-1)])
```

###Create Categorical Variables 
####Urban/Suburban/Rural
####Won by Clinton/Trump in 2016
####Atlanta Metro Area
```{r}
data$old_margin = 100*(ga2016$HRCshare-ga2016$DJTShare)
data$new_margin = (data$BidenVoteShare-data$TrumpVoteShare)
data$margin_change = data$new_margin-data$old_margin

data$past_result[data$old_margin>0] = "Dem"
data$past_result[data$old_margin<0] = "Rep"
data$past_result = as.factor(data$past_result)

data$current_result[data$new_margin>0] = "Dem"
data$current_result[data$new_margin<0] = "Rep"
data$current_result = as.factor(data$current_result)

urbanicity = as.data.frame(read_excel("C:/Users/alber/Downloads/Stat382/Project4/Sub Urb Rural.xlsx"))
data$urbanicity = urbanicity$`Sub/Urb/Rural`
atlanta = c("Fulton", "DeKalb", "Gwinnett", "Cobb", "Clayton", "Cherokee", "Douglas", "Fayette", "Henry", "Rockdale", "Bartow", "Coweta", "Hall", "Forsyth", "Paulding")
data$in_atlanta_metro[data$County %in% atlanta==TRUE] = "Yes"
data$in_atlanta_metro[data$County %in% atlanta==FALSE] = "No"
```
Save Cleaned Data
```{r}
library(writexl)
write_xlsx(data,"C:/Users/alber/Downloads/Stat382/Project4/organized_data.xlsx")
```

#Load cleaned data
```{r}
input = as.data.frame(read_excel("C:/Users/alber/Downloads/Stat382/Project4/Cleaned_Data.xlsx"))
input[,c(24:27)] = lapply(input[,c(24:27)],as.factor)
```


Hypotheses:
1. More educated counties -> more Democratic
2. More educated counties -> shifted more Democratic in the past 4 years
3. Turnout increased more in White counties
4. Higher population density -> more Democratic
5. Higher income counties -> shifted more Democratic in the past 4 years
6. Turnout increase higher in minority counties (<60% White)
7. Older counties are more Republican
#Ask Professor whether to weight the ratios
ORGANIZATION
Variables:
*Education (quantitative) linear association
Income (quantitative) significance test
Urban/Suburban/Rural (categorical) ANOVA
Racial breakdown linear association barplot
Turnout (urban/suburban/rural ANOVA) histogram
Change from 2016 
Population Density 
2016 result by party (categorical)
Share of 65+ or <30

#Exploratory Analysis
```{r}
boxplot(input$margin_change~input$urbanicity)
boxplot(input$new_margin~input$urbanicity)
test <- aov(BidenVoteShare~urbanicity, data = input)
summary(test)
TukeyHSD(test)
```


